---
title: "Nutrient maps"
author: "*Compiled on `r date()` by Jamie Afflerbach"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
  pdf_document:
    toc: true
---

# Summary


# Setup

```{r setup, message = F, warning = F}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)

library(tidyverse)
library(sf)
library(raster)
library(tmap)
library(RColorBrewer)


#point to aquaculture folder on aurora
dir_a <- c('Windows' = '//aurora.nceas.ucsb.edu/aquaculture',
           'Darwin'  = '/Volumes/aquaculture',   
           'Linux'   = '/home/shares/aquaculture')[[ Sys.info()[['sysname']] ]]


#set rainbow color scheme
cols      = colorRampPalette(rev(brewer.pal(9, 'Spectral')))(255) # rainbow color scheme

#grab the worldmap from rnaturalearth
worldmap <- rnaturalearth::ne_download(scale = 110,
                                       type = "countries",
                                       category = "cultural",
                                       destdir = tempdir(),
                                       load = TRUE,
                                       returnclass = "sf")
```

# Data wrangling

## Phosphate data

Load the .csv of Phosphate downloaded to aurora. This data was downloaded from the [World Ocean Atlas 2013](https://www.nodc.noaa.gov/cgi-bin/OC5/woa13/woa13oxnu.pl). We need to skip the first row since it just contains a string.

```{r}
l <- list.files(file.path(dir_a, "input/WOA"), full.names = T)

seasonal <- l[str_detect(l, "an01.csv$")]

phos <- seasonal[str_detect(seasonal, "all_p[1][3-6]")]

f <- function(file){
  
df <- read_csv(file, skip = 1)%>%
  rename(lat = `#COMMA SEPARATED LATITUDE`, 
         long = `LONGITUDE`) %>%
  mutate(season = substr(file, 47, 48))
return(df)
}

phosphate <- map_df(phos,f) %>%
  mutate(season = case_when(
    season == 13 ~ "Winter",
    season == 14 ~ "Spring",
    season == 15 ~ "Summer",
    season == 16 ~ "Fall"
  ))
```

All columns 3 and on contain depth measurements. I'm limiting our range to just the top 10 m.

```{r}
phos_surf <- phosphate %>%
  dplyr::select(1:7,40) %>%
  gather(key = depth, value = phosphate, -lat, -long, -season) %>%
  mutate(depth_m = ifelse(depth == "AND VALUES AT DEPTHS (M):0", 0, depth)) %>%
  dplyr::select(-depth)
```

Let's rasterize phosphate values at 0 m depth (surface)

```{r}
x <- raster(xmn=-180, xmx=180, ymn=-90, ymx=90, res=1, crs="+proj=longlat +datum=WGS84")

#get each depth x season raster

s <- stack()
for(i in seq(0,10,b=5)){
  
  for(j in c("Winter", "Spring", "Summer", "Fall")){
    
    d <- phos_surf %>% filter(depth_m == i,
                              season == j)
    
    p <- rasterize(d[, c('long', 'lat')], x, d[, 'phosphate'], fun=mean)
    names(p) <- paste0("depth_",i,"_",j)
    
    s <- stack(s, p)
  }
  
}
```

Average for each season

```{r, eval = F}

for(i in c("Winter", "Spring", "Summer", "Fall")){

  m <- raster::subset(s, grep(i, names(s), value = T)) %>%
    mean(na.rm=T)
  
  writeRaster(m, filename = paste0("~/github/seaweed/Data/avg_phosphate_", i, ".tif"), overwrite = T)
}

```


## Nitrogen data

Load the .csv of nitrate downloaded to aurora. This data was downloaded from the [World Ocean Atlas 2013](https://www.nodc.noaa.gov/cgi-bin/OC5/woa13/woa13oxnu.pl). We need to skip the first row since it just contains a string.

```{r}

l <- list.files(file.path(dir_a, "input/WOA"), full.names = T)

seasonal <- l[str_detect(l, "an01.csv$")]

nitr <- seasonal[str_detect(seasonal, "all_n[1][3-6]")]

f <- function(file){
  
df <- read_csv(file, skip = 1)%>%
  rename(lat = `#COMMA SEPARATED LATITUDE`, 
         long = `LONGITUDE`) %>%
  mutate(season = substr(file, 47, 48))
return(df)
}

nit <- map_df(nitr,f) %>%
  mutate(season = case_when(
    season == 13 ~ "Winter",
    season == 14 ~ "Spring",
    season == 15 ~ "Summer",
    season == 16 ~ "Fall"
  ))
```

All columns 3 and on contain depth measurements. I'm limiting our range to just the top 10 m.

```{r}
nit_surf <- nit %>%
  dplyr::select(1:7,40) %>%
  gather(key = depth, value = nitrate, -lat, -long, -season) %>%
  mutate(depth_m = ifelse(depth == "AND VALUES AT DEPTHS (M):0", 0, depth)) %>%
  dplyr::select(-depth)
```



Let's rasterize phosphate values at 0 m depth (surface)

```{r}
x <- raster(xmn=-180, xmx=180, ymn=-90, ymx=90, res=1, crs="+proj=longlat +datum=WGS84")

#get each depth x season raster

s <- stack()
for(i in seq(0,10,b=5)){
  
  for(j in c("Winter", "Spring", "Summer", "Fall")){
    
    d <- nit_surf %>% filter(depth_m == i,
                              season  == j)
    
    n <- rasterize(d[, c('long', 'lat')], x, d[, 'nitrate'], fun=mean)
    names(n) <- paste0("depth_",i,"_",j)
    
    s <- stack(s, n)
  }
}
```

Average for each season

```{r, eval = F}

for(i in c("Winter", "Spring", "Summer", "Fall")){

  m <- raster::subset(s, grep(i, names(s), value = T)) %>%
    mean(na.rm=T)
  
  writeRaster(m, filename = paste0("~/github/seaweed/Data/avg_nitrate_", i, ".tif"), overwrite = T)
}

```

Get standard deviation

```{r}
n_sd <- stack(list.files("~/github/seaweed/Data", pattern = "avg_nitrate", full.names = T)) %>%
  calc(fun = sd)

p_sd <- stack(list.files("~/github/seaweed/Data", pattern = "avg_phosphate", full.names = T)) %>%
  calc(fun = sd)
```


### Make maps

Phosphate

```{r, fig.width = 10, fig.height = 6}
p <- list.files("~/github/seaweed/Data", pattern = "avg_phosphate", full.names = T) %>%
  stack()

phosphate_panel <- tm_shape(p) +
  tm_raster(style = "cont", 
            title = "Phosphate (µmol/L)", 
            palette = cols) +
  tm_shape(worldmap)  +
  tm_polygons(col = "gray90") +
  tm_legend(outside = TRUE) +
  tm_layout(legend.title.size = 1, 
            legend.title.fontface = "bold", 
            panel.labels = c("Fall", "Spring", "Summer", "Winter"), 
            panel.label.bg.color = "white",
            legend.text.size = 0.8)

phosphate_panel

tmap_save(phosphate_panel, filename = "~/github/seaweed/Figs/phosphate_seasonal_panel.pdf")
```

Nitrate
```{r, fig.width = 10, fig.height = 6}
n <- list.files("~/github/seaweed/Data", pattern = "avg_nitrate", full.names = T) %>%
  stack()

nitrate_panel <- tm_shape(n) +
  tm_raster(style = "cont", 
            title = "Nitrate (µmol/L)", 
            palette = cols,
            breaks = c(0, 10, 20, 30, 40)) +
  tm_shape(worldmap)  +
  tm_polygons(col = "gray90") +
  tm_legend(outside = TRUE) +
  tm_layout(legend.title.size = 1, 
            legend.title.fontface = "bold", 
            panel.labels = c("Fall", "Spring", "Summer", "Winter"), 
            panel.label.bg.color = "white",
            legend.text.size = 0.8)

tmap_save(nitrate_panel, filename = "~/github/seaweed/Figs/nitrate_seasonal_panel.pdf")
```

standard deviation maps
```{r}
p_sd_map <- tm_shape(p_sd) +
  tm_raster(style = "cont", 
            title = "Phosphate (µmol/L)", 
            palette = cols) +
  tm_shape(worldmap)  +
  tm_polygons(col = "gray90") +
  tm_legend(outside = TRUE,
            main.title = "Standard deviation",
            main.title.position = c("left"),
            main.title.size = 1) +
  tm_layout(legend.title.size = 1, 
            legend.title.fontface = "bold",
            legend.text.size = 0.8,
            title.snap.to.legend = F)

p_sd_map
tmap_save(p_sd_map, filename = "~/github/seaweed/Figs/phosphate_sd.pdf")
```

```{r}
n_sd_map <- tm_shape(n_sd) +
  tm_raster(style = "cont", 
            title = "Nitrate (µmol/L)", 
            palette = cols,
            breaks = seq(0, 17, length.out = 7)) +
  tm_shape(worldmap)  +
  tm_polygons(col = "gray90") +
  tm_legend(outside = TRUE,
            main.title = "Standard deviation",
            main.title.position = c("left"),
            main.title.size = 1) +
  tm_layout(legend.title.size = 1, 
            legend.title.fontface = "bold",
            legend.text.size = 0.8,
            title.snap.to.legend = F)
n_sd_map
tmap_save(n_sd_map, filename = "~/github/seaweed/Figs/nitrate_sd.pdf")
```