---
title: "Explore WOA data"
author: "Jamie Afflerbach"
date: "12/13/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(sf)
library(raster)

#point to aquaculture folder on aurora
dir_a <- c('Windows' = '//aurora.nceas.ucsb.edu/aquaculture',
           'Linux'   = '/home/shares/aquaculture')[[ Sys.info()[['sysname']] ]]

library(RColorBrewer)
cols      = colorRampPalette(brewer.pal(9, 'Spectral'))(255) # rainbow color scheme

library(rworldmap)
#get coarse resolution world from rworldmap
world <- getMap()  
```

Load the .csv of Phosphate downloaded to aurora. This data was downloaded from the [World Ocean Atlas 2013](https://www.nodc.noaa.gov/cgi-bin/OC5/woa13/woa13oxnu.pl). We need to skip the first row since it just contains a string.

```{r}
phos <- read_csv(file.path(dir_a, "input/WOA/woa13_all_p00an01.csv"), skip = 1) %>%
  rename(lat = `#COMMA SEPARATED LATITUDE`, 
         long = `LONGITUDE`)
```

All columns 3 and on contain depth measurements. I'm limiting our range to just the top 50 m.

```{r}
phos_surf <- phos %>%
  dplyr::select(1:13) %>%
  gather(key = depth, value = phosphate, -lat, -long) %>%
  mutate(depth_m = ifelse(depth == "AND VALUES AT DEPTHS (M):0", 0, depth)) %>%
  dplyr::select(-depth)
```

Let's rasterize phosphate values at 0 m depth (surface)

```{r}
zero <- phos_surf %>%
  filter(depth_m == 0)

# in your case (no need to compute rows/colums):
x <- raster(xmn=-180, xmx=180, ymn=-90, ymx=90, res=1, crs="+proj=longlat +datum=WGS84")

phos_ras <- rasterize(zero[, c('long', 'lat')], x, zero[, 'phosphate'], fun=mean)
plot(phos_ras, main = "Phosphate levels (µmol/l) at 0 m (surface)")
```

Trying the same thing from 0 - 30
```{r}
depth_raster <- function(depth){
  
  d <- phos_surf %>% filter(depth_m == depth)
  p <- rasterize(d[, c('long', 'lat')], x, d[, 'phosphate'], fun=mean)
  
  return(p)
}
 
depths <- seq(0, 10, by = 5)

phos <- map(depths, depth_raster) %>% 
  stack() 

min <- range(phos)[[1]]
max <- range(phos)[[2]]

diff <- max-min
plot(diff, main = "Max possible differences in Phosphate between the surface (0 m) and 10 m")
plot(phos)
```

## Nitrogen

Load the .csv of nitrate downloaded to aurora. This data was downloaded from the [World Ocean Atlas 2013](https://www.nodc.noaa.gov/cgi-bin/OC5/woa13/woa13oxnu.pl). We need to skip the first row since it just contains a string.

```{r}
nit <- read_csv(file.path(dir_a, "input/WOA/woa13_all_n00an01.csv"), skip = 1) %>%
  rename(lat = `#COMMA SEPARATED LATITUDE`, 
         long = `LONGITUDE`)
```

All columns 3 and on contain depth measurements. I'm limiting our range to just the top 50 m.

```{r}
nit_surf <- nit %>%
  dplyr::select(1:13) %>%
  gather(key = depth, value = nitrate, -lat, -long) %>%
  mutate(depth_m = ifelse(depth == "AND VALUES AT DEPTHS (M):0", 0, depth)) %>%
  dplyr::select(-depth)
```

```{r}
depth_raster_n <- function(depth){
  
  d <- nit_surf %>% filter(depth_m == depth)
  p <- rasterize(d[, c('long', 'lat')], x, d[, 'nitrate'], fun=mean)
  
  return(p)
}
 
depths <- seq(0, 10, by = 5)

nitrate <- map(depths, depth_raster_n) %>% 
  stack() 

min <- range(nitrate)[[1]]
max <- range(nitrate)[[2]]

diff <- max-min
plot(diff, main = "Max possible differences in Nitrate between the surface (0 m) and 10 m" )
plot(nitrate)
```




Range of N:P ratios for seaweed is 10:1 to 80:1 with optimal 30:1. Due to data resolution, using the range of 10 - 80 still leaves out key areas where we know marine plants/seaweed are growing naturally. Specifically the west coast of N and S America. Since the data resolution is so coarse, we increased the range slightly until we were capturing those locations.


```{r}
#average values from 0 - 10 m

avg_p <- mean(phos)
avg_n <- mean(nitrate)

n_to_p <- avg_n/avg_p

plot(n_to_p, col = cols)
hist(n_to_p)

```

select between 4 and 80

```{r}
n_to_p[n_to_p < 4] <- NA
n_to_p[n_to_p > 80] <- NA

plot(n_to_p, col = cols)

```

Now we want to cut out all high seas areas and only keep feasible areas in EEZs.

```{r}
eezs <- read_sf(file.path(dir_a, "input/world_eezs/regions_2017_update.shp"), quiet = T) %>%
  filter(rgn_type == "eez",
         rgn_name != "Antarctica") %>%
  st_transform(crs = 4326) %>%
  as('Spatial')

n_to_p_eez <- mask(n_to_p, eezs)

```


Get countries where production is currently happening

```{r}

fao <- read_csv("~/github/seaweed/Data/FAO_AquaticPlants_1950-2016.csv") %>%
  filter(Environment == "Marine") %>%
  gather(key = year, value = tons, -`Land Area`, -`Ocean Area`, -Environment, -Species, -`Scientific name`) %>%
  filter(!str_detect(year, "S_"),
         year > 2010,
         tons > 0)

ctrys <- unique(fao$`Land Area`)

worldmap <- rnaturalearth::ne_download(scale = 110,
                                       type = "countries",
                                       category = "cultural",
                                       destdir = tempdir(),
                                       load = TRUE,
                                       returnclass = "sf")

#add aquaculture countries

worldmap_aq <- worldmap %>%
  select(NAME) %>%
  mutate(aq_production = ifelse(NAME %in% ctrys, 1, 0))

sum(worldmap_aq$aq_production)
```

Looks like we only have 26 of the 37 countries identified. Probably due to name mismatches. Let's see what those are

```{r}

c <- worldmap_aq %>% filter(aq_production == 1) %>% .$NAME

setdiff(ctrys, c)
```

```{r}
aq_map <- worldmap_aq %>%
  mutate(aq_production = ifelse(NAME %in% c("Russia", "South Korea", "North Korea", "Vietnam", "Tanzania", "Taiwan", "Solomon Is."), 1, aq_production)) %>%
  as('Spatial')

#this world map doesn't have Cook Islands, Kiribati, Saint Lucia or Zanzibar
```

Plot N:P plus countries producing marine seaweed
```{r}

land_cols <- c("beige", "darkolivegreen4")

tmap_mode("plot")

tm_shape(n_to_p_eez) +
  tm_raster(palette = "-RdYlBu", style = "cont", auto.palette.mapping=FALSE, title = "N:P") +
  tm_shape(aq_map) +
  tm_polygons("aq_production", 
              legend.title = "Countries currently farming seaweeds", 
              palette = land_cols, alpha = 0.7,  legend.show = F) +
tm_legend(outside = TRUE)


```





