---
title: "Hypoxic, Eutrophic & OA map"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
  pdf_document:
    toc: true
---

## Summary

This script creates a map showing regions of the world where acidic conditions exist (aragonite saturation state <= 3), along with specific sites identified as either Eutrophic or Hypoxic. At the end of this script, multiple statistics are run to glean information about countries that are suitable for seaweed farming. For example, how many countries that have been identified as suitable for seaweed farming are also experiencing hypoxic, eutrophic AND acidic conditions? (Hint it's 24).

## Setup

```{r setup, message = F, warning = F}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)

library(tidyverse)
library(sf)
library(raster)
library(tmap)
library(RColorBrewer)
library(readxl)

#point to aquaculture folder on aurora
dir_a <- c('Windows' = '//aurora.nceas.ucsb.edu/aquaculture',
           'Darwin'  = '/Volumes/aquaculture',    
           'Linux'   = '/home/shares/aquaculture')[[ Sys.info()[['sysname']] ]]

#set rainbow color scheme
cols      = colorRampPalette(brewer.pal(9, 'Spectral'))(255) # rainbow color scheme
```


## Data cleaning

### Hypoxic/eutrophic data

Read in hypoxic eutrophic site data

```{r}
sites <- read_excel("~/github/seaweed/Data/wri_eutrophic_hypoxic_dataset_2011-03.xls")

#filter for only eutrophic and hypoxic sites
sites_he <- sites %>%
  mutate(category = ifelse(Classification == "eutrophic", "Eutrophic", Classification)) %>%
  filter(!category %in% c("Improved", "improved"))
```

Create `sf` object from sites

```{r}
sites_sf <- st_as_sf(sites_he,coords = c("Long", "Lat"), crs = 4326) %>%
  st_transform(crs = "+proj=moll +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +units=m +no_defs")
```

### Ocean Acidification

Load OA raster, interpolate to gapfill missing cells especially at the Arctic, and then remove all cells with Aragonite saturation state greater than 3

```{r}
#load OA raster from 2016
oa    <- raster(file.path(dir_a, "ARF/oa/global_arag_avg_moll_2016.tif")) 
ocean <- raster(file.path(dir_a, "ARF/oa/ocean.tif")) %>% #ocean raster at 1km touse for mask
           resample(oa)
  
#interpolate across NA
xy <- data.frame(xyFromCell(oa, 1:ncell(oa)))                         #get xy coords into dataframe
v  <- getValues(oa)                                                   #get cell values 
tmpdf <- cbind(xy, v)%>%filter(!is.na(v))                             #create dataframe of x,y, and values. remove NAs (throws error since these are cells we are interpolating over)
mg <- gstat::gstat(id = "v", formula = v~1, locations = ~x+y, data=tmpdf,
            nmax=7, set=list(idp = 2)) #define model. power function = 2, this is default for idw models
oa_int <- interpolate(oa, mg, progress='text')  
oa_mask <- mask(oa_int, ocean)

oa_mask[oa_mask>3.05]<-NA #using 3.05 to get contour lines for Aragonite values at 3
plot(oa_mask, col = cols)
```

## Make map

### Get world map

```{r}
#grab the worldmap from rnaturalearth
worldmap <- rnaturalearth::ne_download(scale = 110,
                                       type = "countries",
                                       category = "cultural",
                                       destdir = tempdir(),
                                       load = TRUE,
                                       returnclass = "sf")
```

### Final map and save

```{r}
#make map
hyp_eut_map <- 
  tm_shape(oa_mask) +
  tm_raster(palette = "RdYlBu", style = "cont", title = "Ω (aragonite \nsaturation state)") +
  # tm_shape(oa_cont) +
  # tm_lines() +
  tm_shape(worldmap) +
  tm_polygons(col = "ivory", lwd = 0.5, alpha = 0.85) +
  tm_shape(sites_sf) +
  tm_symbols(col = "category", border.col = "black", border.lwd = 0.3, 
             size = 0.05, alpha = 0.7, title.col = " ", palette = "PRGn", legend.col.show = FALSE) +
  tm_legend(outside = TRUE) +
  tm_add_legend(type = "symbol", shape = 21, labels = "Hypoxic site", col = "#AF8DC3", size = 0.4) +
  tm_add_legend(type = "symbol", shape = 21, labels = "Eutrophic site", col = "#7FBF7B", size = 0.4) +
  tm_layout(legend.title.size = 0.8, legend.title.fontface = "bold", frame = F) 


tmap_save(hyp_eut_map, filename = "../../Figs/hyp_eut_map.png", width = 6, height = 2.5, units = "in")

hyp_eut_map
```


## Get country specific statistics

Get list of countries that have hypoxic, eutrophic and/or acidic conditions. 

```{r}
#suitable countries for seaweed
ctrys <- read_csv("~/github/seaweed/Data/sw_feasible_countries.csv")

#reading in the OHI EEZ shapefile from aurora and filtering just for EEZs while removing Antarctica
eezs <- read_sf(file.path(dir_a, "input/world_eezs/regions_2017_update.shp"), quiet = T) 

#intersect the sites with eezs
st_crs(sites_sf) <- "+proj=moll +lon_0=0 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs"
sites_int <- st_intersection(eezs, sites_sf)

#get the columns we want (country, classification, site info)
si <- sites_int %>%
  dplyr::select(rgn_name, System, Country, Classification) %>%
  distinct()
#remove geometry
st_geometry(si) <- NULL

#dataframe that lists each country, and how many sites are eutrophic or hypoxic within each, filtered for feasible countries
ctrys_w_eut_hyp <- si %>%
  mutate(Classification = ifelse(Classification == "eutrophic", "Eutrophic", Classification)) %>%
  group_by(rgn_name) %>%
  add_count(Classification) %>%
  dplyr::select(rgn_name, Classification, n) %>%
  distinct() %>%
  filter(rgn_name %in% ctrys$country)
```


Get list of countries that have acidic conditions (aragonite saturation state < 3)

```{r}
#overlay EEZ shapefile on our masked out OA layer to get countries with cells w/ acidic conditions
oa_ctrys <- extract(oa_mask, eezs, df = TRUE)

#remove NA cells and match up IDs
oa_ctry <- oa_ctrys %>%
  filter(!is.na(x)) 

#grab IDs from the extract data frame
ctry_ids <- unique(oa_ctry$ID)

#create a column in eezs that is equivalent to the rownames (which are the IDs returned from extract)
eezs$IDs <- row.names(eezs)

#see what countries those are from eezs
a <- eezs %>%
  filter(IDs %in% ctry_ids)
st_geometry(a) <- NULL

#get final list of countries w/ acidic conditions
final_oa_countries <- a %>%
  filter(rgn_type != "fao") %>%
  dplyr::select(rgn_name) %>%
  distinct() %>%
  filter(rgn_name %in% ctrys$country)
```

Now we combine the hypoxic/eutrophic dataframe with the ocean acidification one

```{r}
b <- ctrys_w_eut_hyp %>%
  bind_rows(final_oa_countries) %>% #add in the OA countries
  mutate(Classification = ifelse(is.na(Classification), "oa", Classification),
         n = ifelse(Classification == "oa", 1, n)) %>%
  spread(key = Classification, value = n) %>%
  mutate(OA = ifelse(rgn_name %in% final_oa_countries$rgn_name, 1, 0),
         Eutrophic = ifelse(!is.na(Eutrophic), 1, 0),
         Hypoxic = ifelse(!is.na(Hypoxic), 1, 0)) %>%
  dplyr::select(-oa) %>%
  mutate(sum = sum(OA, Eutrophic, Hypoxic))

write.csv(b, file = "~/github/seaweed/Data/countries_w_hyp_eut_oa_conditions.csv")
```


Time to run some stats.

How many countries have acidic conditions?

```{r}
b %>%
  filter(OA == 1) %>%
  nrow()
```
And # of suitable countries with eutrophic, hypoxic, OR acidic conditions, please! 

These countries have all three conditions

```{r}
b %>% filter(sum == 3) %>% .$rgn_name
```

Number of feasible countries with Eutrophic conditions
```{r}
b %>% filter(Eutrophic==1) %>% nrow()
```


Number of feasible countries with Hypoxic conditions
```{r}
b %>% filter(Hypoxic == 1) %>% nrow()
```

Number of feasible countries with Eutrophic and Hypoxic
```{r}
b %>% filter(Eutrophic == 1 & Hypoxic == 1) %>% nrow()
```







